Tuning for Software Analytics: is it Really Necessary?

### 2016-04-26 tuningpredictor.tex
Updates from ShareLaTeX

Are the impacts of tuning addressed in the defect prediction literature?
To answer that question,  in Jan 2016 we searched scholar.google.com for the conjunction of ``data mining" and ``software engineering" and  ``defect prediction"\footnote{More details can be found at https://goo.gl/Inl9nF}.
After sorting by the citation count and discarding the non-SE papers (and those without a pdf link), we read over this sample
of  50 highly-cited SE defect prediction papers. 
What we found in that sample was that few authors
acknowledged the impact of tunings (exceptions:~\cite{Gao:2011,lessmann2008benchmarking}).
Overall,  80\% of papers in our sample {\em did not} adjust
the ``off-the-shelf'' configuration of the data miner (e.g.~\cite{me07b,Moser:2008,Elish2008649}). Of the remaining papers:
\bi
\item
Some papers in our sample  explored   data super-sampling~\cite{4271036} or data sub-sampling techniques via  automatic methods (e.g. ~\cite{Gao:2011,me07b,4271036,Kim:2011}) 
or via some domain principles (e.g. ~\cite{Moser:2008,Nagappan:2008,Hassan:2009}).
As an example of the latter, Nagappan et al.~\cite{Nagappan:2008} checked if metrics related to organizational structure were relatively more powerful for predicting software defects. 
However, it should be noted that  these studies varied the input data but
not the   ``off-the-shelf''   settings of the data miner.
\item
A few other papers did acknowledge that one data miner may not be appropriate
for all data sets.  Those papers tested  different  
``off-the-shelf'' data miners on the same data set.
For example, Elish et al.\cite{Elish2008649}  compared support vector
machines to other data miners for the purposes of defect prediction. SVM's execute via a ``kernel function'' which should be specially selected for different data sets and
the Elish et al. paper  makes no mention of any SVM tuning study.  
To be fair to Elish et al., we hasten to add that we
ourselves have  published
papers using ``off-the-shelf'' tunings~\cite{me07b} since,
prior to this paper it was unclear to us how to effectively
navigate the large space of possible tunings.
\ei
Over our entire sample, there was only  one paper that conducted a somewhat extensive tuning study.
Lessmann et al.\cite{lessmann2008benchmarking} tuned parameters for some of their algorithms using  a {\em grid search}; i.e. divide all $C$ configuration
options into $N$ values, then try all   $N^C$ combinations.
This is a slow approach-- we have explored grid search for 
defect prediction and found it takes days to terminate~\cite{me07b}.
Not only that, we found that grid search can miss
important optimizations~\cite{baker07}.
Every grid has ``gaps'' between each grid division which means
that a supposedly rigorous grid search can still miss
important configurations~\cite{Bergstra2012}. 
Bergstra and Bengio~\cite{Bergstra2012} comment that for most data sets only a few of the tuning parameters really matter-- which means that
much of the runtime associated with grid search is actually wasted.
Worse still, Bergstra and Bengio  comment that 
the 
important tunings are   different   for different
data sets-- a 
 phenomenon that makes grid search a poor choice for configuring data mining
 algorithms for new data sets.

@@

This section discusses defect prediction,
which is the particular
task explored by our optimizers.
Note that this section repeats much of 
our standard introduction to defect prediction~\cite{me15:book1},
as well as presenting    some new results from Rahman et al.~\cite{rahman14:icse}.

\input{dm101}

@@

\begin{equation}\label{eq:yesno}\scriptsize
\mathit{inspect}= \left\{
\begin{array}{ll}
d_i \ge T \rightarrow \mathit{Yes}\\
d_i <   T \rightarrow \mathit{No} ,
\end{array}\right.
\end{equation}
The splitting process is controlled by numerous tuning parameters.
If data contains more than {\em min\_sample\_split}, then a split is attempted.
On the other hand, if a split contains no more than {\em min\_samples\_leaf}, then the recursion stops. CART and Random Forest use a 
user-supplied constant for this parameter while
WHERE-based learner firstly computes this parameter $m$={\em min\_samples\_leaf} from the size of the data
sets via  $m=\mathit{size}^\mathit{min\_size}$ to build an initial  clustering tree.
Note that WHERE builds {\em two} trees: the initial clustering tree (to find similar sets of data)
then a final decision tree (to learn rules that predict for each similar cluster)\footnote{A
frequently asked question is why does WHERE build two trees--
would not   a single tree suffice? The answer is, as shown below,  tuned WHERE's twin-tree approach 
generates very precise predictors.}.
The tuning parameter  {\em min\_sample\_ split } controls the construction of the
final decision tree (so, for WHERE-based learner,
{\em min\_size} and {\em min\_sample\_split} are the parameters to be tuned).

These learners use different techniques to explore the splits:
\bi
\item
CART finds the attributes whose ranges contain rows with least variance in the number
of defects\footnote{If an attribute ranges $r_i$ is found in 
$n_i$ rows each with a  defect count variance of $v_i$, then CART seeks the attributes
whose ranges minimizes $\sum_i \left(\sqrt{v_i}\times n_i/(\sum_i n_i)\right)$.}.
\item
Random Forest    divides data like CART then  builds $F>1$  trees,
each time using some random subset of
the attributes. 
\item
When building the initial cluster tree, WHERE projects the data on to a dimension it synthesizes from the raw data using
a process analogous to principle component analysis~\cite{jolliffe2002principal}

@@

\input{algo}

@@

Our defect data comes from the PROMISE repository \footnote{http://openscience.us/repo}
and pertains to 
open source Java systems defined in terms of \tab{ck}:  {\it ant}, {\it camel}, {\it ivy}, {\it jedit}, {\it log4j}, {\it lucene},
{\it poi}, {\it synapse}, {\it velocity} and {\it xerces}.

@@

\subsection{RQ2:  Does Tuning Change a Learner's Ranking ?}\label{sect:rank}
Researchers often use performance criteria to assert that one learner is better than 
another~\cite{lessmann2008benchmarking,me07b,hall11}. For example:
\be
\item
Lessmann et al.~\cite{lessmann2008benchmarking} conclude that
Random Forest is considered to be statistically 
better than CART. 
\item
Also, in Hall et al.'s   systematic literature review\cite{hall11}, it is argued
that defect predictors based on simple 
modeling techniques such as LR perform better than complicated techniques such as Random Forest\footnote{By three measures,
Random Forest
is more complicated than LR. Firstly, LR builds one model
while Random Forest builds many models. Secondly, LR is just
a model construction tool while Random Forest needs both
a tool to construct its forest {\em and} a second tool
to  infer some conclusion from all the members of that forest.
Thirdly, the LR model can be printed in a few lines while the multiple
models learned by Random
Forest model would take up multiple pages of output.}.
\ee
Given tuning, how stable are these  conclusions?
Before answering this issue, we digress for two comments.

@@

%%%%time for prec %%%%%%
\begin{table*}[!ht]
\renewcommand{\baselinestretch}{0.75}
\scriptsize
\centering
  \begin{tabular}{r|c |c |c |c |c |c }
    Datasets & Tuned\_Where & Naive\_Where & Tuned\_CART & Naive\_CART & Tuned\_RanFst & Naive\_RanFst\\
    \hline
    antV0 & 50 / 95.47 & 1.65 & 60 / 5.08 & 0.08 & 60 / 9.78 & 0.20\\
    antV1 & 60 / 224.67 & 3.03 & 50 / 6.52 & 0.12 & 60 / 14.13 & 0.25\\
    antV2 & 70 / 644.99 & 8.24 & 50 / 9.00 & 0.24 & 60 / 16.75 & 0.44\\
    camelV0 & 70 / 690.62 & 7.93 & 70 / 12.68 & 0.24 & 110 / 28.49 & 0.34\\
    camelV1 & 60 / 1596.77 & 23.56 & 60 / 17.13 & 0.27 & 70 / 33.96 & 0.77\\
    ivy & 60 / 66.69 & 0.97 & 60 / 4.26 & 0.07 & 60 / 8.89 & 0.19\\
    jeditV0 & 80 / 459.30 & 5.33 & 80 / 8.69 & 0.11 & 90 / 18.40 & 0.32\\
    jeditV1 & 60 / 421.56 & 6.59 & 80 / 9.05 & 0.12 & 80 / 17.93 & 0.36\\
    jeditV2 & 90 / 595.56 & 6.88 & 60 / 7.90 & 0.14 & 110 / 27.34 & 0.38\\
    log4j & 50 / 76.09 & 1.33 & 50 / 2.60 & 0.06 & 80 / 9.69 & 0.15\\
    lucene & 80 / 236.45 & 2.60 & 70 / 6.07 & 0.10 & 60 / 9.77 & 0.25\\
    poiV0 & 60 / 263.12 & 3.92 & 70 / 7.42 & 0.09 & 130 / 25.86 & 0.28\\
    poiV1 & 50 / 398.33 & 6.94 & 70 / 9.31 & 0.13 & 50 / 12.67 & 0.29\\
    synapse & 70 / 144.09 & 1.85 & 50 / 3.88 & 0.07 & 50 / 8.13 & 0.19\\
    velocity & 60 / 184.10 & 2.68 & 50 / 4.27 & 0.07 & 100 / 15.18 & 0.21\\
    xercesV0 & 60 / 136.87 & 1.98 & 80 / 9.17 & 0.10 & 70 / 14.17 & 0.22\\
    xercesV1 & 80 / 1173.92 & 12.78 & 60 / 10.47 & 0.16 & 50 / 18.27 & 0.40\\
  \end{tabular}
  \caption{Evaluations/runtime and runtime for tuned and default learners(in sec), optimizing for precision.}
  \label{tab:etimes}
\end{table*}

%%%%time for F %%%%%%
\begin{table*}[!ht]
\renewcommand{\baselinestretch}{0.75}
\scriptsize
\centering
  \begin{tabular}{r|c |c |c |c |c |c }
    Datasets & Tuned\_Where & Naive\_Where & Tuned\_CART & Naive\_CART & Tuned\_RanFst & Naive\_RanFst\\
    \hline
    antV0 & 50 / 93.38 & 1.39 & 50 / 3.52 & 0.08 & 70 / 9.89 & 0.17\\
    antV1 & 60 / 186.95 & 3.18 & 50 / 6.18 & 0.12 & 60 / 13.39 & 0.25\\
    antV2 & 90 / 654.34 & 8.08 & 60 / 8.79 & 0.18 & 120 / 27.56 & 0.36\\
    camel & 50 / 543.28 & 9.65 & 80 / 17.00 & 0.28 & 70 / 22.52 & 0.41\\
    camelV1 & 60 / 1808.03 & 26.98 & 110 / 31.92 & 0.28 & 70 / 37.00 & 0.85\\
    ivy & 60 / 74.50 & 1.18 & 60 / 4.72 & 0.08 & 60 / 10.39 & 0.21\\
    jeditV0 & 80 / 518.47 & 6.11 & 60 / 7.9
 & 0.10 & 60 / 14.32 & 0.37\\
    jeditV1 & 70 / 576.29 & 6.89 & 70 / 8.13 & 0.10 & 70 / 17.42 & 0.34\\
    jeditV2 & 80 / 657.59 & 7.93 & 70 / 10.34 & 0.15 & 80 / 20.20 & 0.40\\
    log4j & 70 / 123.48 & 1.59 & 50 / 2.92 & 0.08 & 50 / 7.67 & 0.17\\
    lucene & 60 / 219.02 & 3.68 & 60 / 6.89 & 0.12 & 70 / 13.06 & 0.35\\
    poiV0 & 60 / 314.53 & 4.82 & 60 / 7.80 & 0.10 & 80 / 19.29 & 0.32\\
    poiV1 & 50 / 446.05 & 7.55 & 50 / 7.62 & 0.14 & 110 / 27.23 & 0.36\\
    synapse & 60 / 138.75 & 1.83 & 60 / 4.87 & 0.08 & 90 / 13.29 & 0.17\\
    velocity & 60 / 211.88 & 3.13 & 60 / 5.51 & 0.10 & 60 / 11.58 & 0.27\\
    xercesV0 & 80 / 178.49 & 2.02 & 60 / 7.47 & 0.11 & 80 / 17.31 & 0.28\\
    xercesV1 & 80 / 1370.89 & 14.42 & 60 / 11.07 & 0.19 & 80 / 25.27 & 0.46\\
  \end{tabular}
  \caption{Evaluations/runtime and runtime for tuned and default learners(in sec), optimizing for F-Measure.}
\label{tab:ftimes}
\end{table*}

@@

The number of evaluations/runtime used by our optimizers is shown in \tab{etimes}
and \tab{ftimes}.
WHERE's runtime are slower than CART and Random Forest since WHERE has yet to benefit from decades
of implementation experience with these older algorithms. For example, SciKitLearn's  CART and Random Forest
 make extensive use of an underlying C library whereas WHERE is a purely interpreted Python.

Looking over \tab{etimes} and \tab{ftimes}, the general pattern is that 50 to 80 evaluations suffice for finding the tuning
improvements reported in this paper. 
50 to 80 evaluations are  much fewer than our pre-experimental intuition.
Prior to this paper, the authors have conducted numerous explorations of evolutionary algorithms
for search-based SE applications~\cite{krall15,krall15:hm,fea02a,me07f,Green}. Based
on that work, our expectations were that non-parametric evolutionary optimization would
take thousands, if not millions, of evaluations of candidate tunings. This turned out not
to be that case.

@@

 In \fig{features}, we show how tuning selects the optimal values for tuned parameters. For space limitation, only four parameters from WHERE learner are selected as representatives and all the others can be found in our online support documents\footnote{\url{https://goo.gl/aHQKtU}}.
 Note that
 the tunings learned were different in different data sets and for different goals.
Also, the tunings learned by DE
were often very different to the default (the default values for {\em threshold}, {\em infoPrune}, {\em min\_Size} and {\em wriggle} are $0.5$, $0.33$, $0.5$ and $0.2$, respectively). That is, to achieve the performance improvements seen in the paper,
the default tuning parameters required a wide range of adjustments.

@@

\balance

>>>>

Are the impacts of tuning addressed in the defect prediction literature?
To answer that question,  in Jan 2016 we searched scholar.google.com for the conjunction of ``data mining" and ``software engineering" and  ``defect prediction"
(more details can be found at https://goo.gl/Inl9nF).
After sorting by the citation count and discarding the non-SE papers (and those without a pdf link), we read over this sample
of  50 highly-cited SE defect prediction papers. 
What we found in that sample was that few authors
acknowledged the impact of tunings (exceptions:~\cite{Gao:2011,lessmann2008benchmarking}).
Overall,  80\% of papers in our sample {\em did not} adjust
the ``off-the-shelf'' configuration of the data miner (e.g.~\cite{me07b,Moser:2008,Elish2008649}). Of the remaining papers:
\bi
\item
Some papers in our sample  explored   data super-sampling~\cite{4271036} or data sub-sampling techniques via  automatic methods (e.g. ~\cite{Gao:2011,me07b,4271036,Kim:2011}) 
or via some domain principles (e.g. ~\cite{Moser:2008,Nagappan:2008,Hassan:2009}).
As an example of the latter, Nagappan et al.~\cite{Nagappan:2008} checked if metrics related to organizational structure were relatively more powerful for predicting software defects. 
However, it should be noted that  these studies varied the input data but
not the   ``off-the-shelf''   settings of the data miner.
\item
A few other papers did acknowledge that one data miner may not be appropriate
for all data sets.  Those papers tested  different  
``off-the-shelf'' data miners on the same data set.
For example, Elish et al.\cite{Elish2008649}  compared support vector
machines to other data miners for the purposes of defect prediction. SVM's execute via a ``kernel function'' which should be specially selected for different data sets and
the Elish et al. paper  makes no mention of any SVM tuning study.  
To be fair to Elish et al., we hasten to add that we
ourselves have  published
papers using ``off-the-shelf'' tunings~\cite{me07b} since,
prior to this paper it was unclear to us how to effectively
navigate the large space of possible tunings.
\ei
Over our entire sample, there was only  one paper that conducted a somewhat extensive tuning study.
Lessmann et al.\cite{lessmann2008benchmarking} tuned parameters for some of their algorithms using  a {\em grid search}; i.e. divide all $C$ configuration
options into $N$ values, then try all   $N^C$ combinations.
This is a slow approach-- we have explored grid search for 
defect prediction and found it takes days to terminate~\cite{me07b}.
Not only that, we found that grid search can miss
important optimizations~\cite{baker07}.
Every grid has ``gaps'' between each grid division which means
that a supposedly rigorous grid search can still miss
important configurations~\cite{Bergstra2012}. 
Bergstra and Bengio~\cite{Bergstra2012} comment that for most data sets only a few of the tuning parameters really matter-- which means that
much of the runtime associated with grid search is actually wasted.
Worse still, Bergstra and Bengio  comment that 
the 
important tunings are   different   for different
data sets-- a 
 phenomenon that makes grid search a poor choice for configuring data mining
 algorithms for new data sets.

@@

This section discusses defect prediction,
which is the particular
task explored by our optimizers.

Human programmers are clever, but flawed. Coding  adds functionality, but also defects.
Hence, software sometimes crashes (perhaps at the most awkward or dangerous moment) or delivers
the wrong functionality. For a very long list of software-related errors,
see  Peter Neumann's ``Risk Digest'' at catless.ncl.ac.uk/Risks.

Since programming inherently
introduces defects into  programs, it is important to test them before they're used.
Testing is expensive.
Software assessment budgets are finite
while assessment effectiveness increases 
exponentially with assessment effort.
For example, for  black-box testing methods,
a {\em linear} increase
in the confidence $C$ of finding  defects
can take {\em exponentially} more effort:
\bi
\item
A randomly selected 
input to a program will find a fault with probability $p$.
\item
After $N$ random black-box tests, the chances of the inputs 
not revealing any fault 
is $(1-p)^N$. 
\item
Hence, the chances $C$ of seeing the fault is $1-(1-p)^N$
which can be rearranged to 
 $N(C,p)=log(1 -
C)/log(1-p)$. 
\item
For example, $N(0.90,10^{-3})=2301$
but $N(0.98,10^{-3})=3901$; i.e. nearly double the number of tests.
\ei
Exponential costs quickly exhaust finite resources so
standard practice is to apply the best
available  methods on code sections that seem   most critical. 
But 
any method that focuses on parts of the code
can blind us to defects in other areas. Some  {\em
lightweight sampling policy} should be used to explore the rest of the system.  This
sampling policy will always be incomplete.
Nevertheless, it is the only option when
resources prevent a complete assessment of everything.

One such lightweight sampling policy is defect predictors learned from static code attributes.
Given software described in the attributes of \tab{ck},   data miners can
learn where the probability of software defects is highest.

The rest of this section argues that such defect predictors are   {\em easy to
use}, {\em widely-used}, and {\em useful} to use.

\begin{table*}[t]
\renewcommand{\baselinestretch}{0.8}\begin{center}
{\scriptsize
\begin{tabular}{c|l|p{4.7in}}
amc & average method complexity & e.g. number of JAVA byte codes\\\hline
avg\_cc & average McCabe & average McCabe's cyclomatic complexity seen
in class\\\hline
ca & afferent couplings & how many other classes use the specific
class. \\\hline
cam & cohesion amongst classes & summation of number of different
types of method parameters in every method divided by a multiplication
of number of different method parameter types in whole class and
number of methods. \\\hline
cbm &coupling between methods &  total number of new/redefined methods
to which all the inherited methods are coupled\\\hline
cbo & coupling between objects & increased when the methods of one
class access services of another.\\\hline
ce & efferent couplings & how many other classes is used by the
specific class. \\\hline
dam & data access & ratio of the number of private (protected)
attributes to the total number of attributes\\\hline
dit & depth of inheritance tree &\\\hline
ic & inheritance coupling &  number of parent classes to which a given
class is coupled (includes counts of methods and variables inherited)
\\\hline
lcom & lack of cohesion in methods &number of pairs of methods that do
not share a reference to an instance variable.\\\hline
locm3 & another lack of cohesion measure & if $m,a$ are  the number of
$methods,attributes$
in a class number and $\mu(a)$  is the number of methods accessing an
attribute, 
then
$lcom3=((\frac{1}{a} \sum_j^a \mu(a_j)) - m)/ (1-m)$.
\\\hline
loc & lines of code &\\\hline
max\_cc & maximum McCabe & maximum McCabe's cyclomatic complexity seen
in class\\\hline
mfa & functional abstraction & number of methods inherited by a class
plus number of methods accessible by member methods of the
class\\\hline
moa &  aggregation &  count of the number of data declarations (class
fields) whose types are user defined classes\\\hline
noc &  number of children &\\\hline
npm & number of public methods & \\\hline
rfc & response for a class &number of  methods invoked in response to
a message to the object.\\\hline
wmc & weighted methods per class &\\\hline
\rowcolor{lightgray}
defect & defect & Boolean: where defects found in post-release bug-tracking systems.
\end{tabular}
}
\end{center}
\caption{OO measures used in our defect data sets.}\label{tab:ck}
\end{table*}

{\em Easy to use:} Static code attributes can be automatically collected, even for very large systems~\cite{nagappan05}.
Other methods, like  manual code reviews, are far slower and far more labor-intensive.
For example, depending on the review methods, 8 to 20 LOC/minute can be
inspected and this effort repeats for all members of the review team,
which can be as large as four or six people~\cite{me02f}. 
{\em Widely used:}  Researchers and industrial practitioners  use static attributes to guide software 
quality predictions.
 Defect prediction models have been reported
  at Google~\cite{lewis13}.
Verification and validation (V\&V) textbooks
\cite{rakitin01} advise using static code complexity attributes
to decide which modules are worth manual inspections.

{\em Useful:}
Defect predictors often  find the location of  70\% (or more)
of the defects in code~\cite{me07b}.
Defect predictors have some level of generality:
predictors learned at NASA~\cite{me07b} have also been found useful elsewhere
(e.g. in Turkey~\cite{tosun10,tosun09}).
The success of this method in  predictors in finding bugs is   markedly
higher than other currently-used
industrial
methods such as manual code reviews. For example, 
a  panel at {\em IEEE Metrics
2002}~\cite{shu02} concluded that manual software  reviews can find ${\approx}60\%$ 
of defects.
In another work, 
Raffo documents the typical    defect detection capability of
industrial review methods:   around 50\%
 for full Fagan inspections~\cite{fagan76} to
21\% for less-structured inspections.

Not only do static code defect predictors perform well compared to manual methods,
they also are competitive with certain automatic methods.
A recent study at ICSE'14, Rahman et al.~\cite{rahman14:icse} compared
(a) static code analysis tools FindBugs, Jlint, and Pmd and (b)
static code defect predictors
(which they called ``statistical defect prediction'') built using logistic regression.
They found  no significant differences in the cost-effectiveness
of these  approaches. Given this equivalence, it is significant to note that 
static code defect prediction can be quickly adapted to new languages by building lightweight
parsers that find   information like \tab{ck}. The same is not true for   static code analyzers-- these need  extensive modification before they can be used on new
languages.

@@

\begin{equation}\label{eq:yesno}\scriptsize
\mathit{inspect}= \left\{
\begin{array}{ll}
d_i \ge T \rightarrow \mathit{Yes}\\
d_i <   T \rightarrow \mathit{No} ,
\end{array}\right.
\end{equation}
where $d_i$ is the number of observed issues and $T$
is some threshold defined by an engineering judgement; we use $T=1$.

The splitting process is controlled by numerous tuning parameters.
If data contains more than {\em min\_sample\_split}, then a split is attempted.
On the other hand, if a split contains no more than {\em min\_samples\_leaf}, then the recursion stops. CART and Random Forest use a 
user-supplied constant for this parameter while
WHERE-based learner firstly computes this parameter $m$={\em min\_samples\_leaf} from the size of the data
sets via  $m=\mathit{size}^\mathit{min\_size}$ to build an initial  clustering tree.
Note that WHERE builds {\em two} trees: the initial clustering tree (to find similar sets of data)
then a final decision tree (to learn rules that predict for each similar cluster).
A
frequently asked question is why does WHERE build two trees--
would not   a single tree suffice? The answer is, as shown below,  tuned WHERE's twin-tree approach 
generates very precise predictors.
As to the rest of WHERE's parameters, the 
 parameter  {\em min\_sample\_ split } controls the construction of the
final decision tree (so, for WHERE-based learner,
{\em min\_size} and {\em min\_sample\_split} are the parameters to be tuned).

These learners use different techniques to explore the splits:
\bi
\item
CART finds the attributes whose ranges contain rows with least variance in the number
of defects. If an attribute ranges $r_i$ is found in 
$n_i$ rows each with a  defect count variance of $v_i$, then CART seeks the attributes
whose ranges minimizes $\sum_i \left(\sqrt{v_i}\times n_i/(\sum_i n_i)\right)$.
\item
Random Forest    divides data like CART then  builds $F>1$  trees,
each time using some random subset of
the attributes. 
\item
When building the initial cluster tree, WHERE projects the data on to a dimension it synthesizes from the raw data using
a process analogous to principle component analysis~\cite{jolliffe2002principal}

@@

\begin{algorithm}[!t]

\scriptsize
\begin{algorithmic}[1]
\Require $\mathit{np} = 10$, $f=0.75$, $cr=0.3$, $\mathit{life} = 5$, $\mathit{Goal} \in \{\mathit{pd},f,...\}$
\Ensure $S_{best}$
\vspace{2mm}
\Function{DE}{$\mathit{np}$, $f$, $cr$, $\mathit{life}$, $\mathit{Goal}$}
 \State $Population  \gets $ InitializePopulation($\mathit{np}$)   
 \State $S_{best} \gets $GetBestSolution($Population $)
 \While{$\mathit{life} > 0$}
\State $NewGeneration \gets \emptyset$
\For{$i=0 \to \mathit{np}-1$}
\State $S_i \gets$ Extrapolate($Population [i], Population , cr, f$)
\If {Score($S_i$) >Score($Population [i]$)}
\State $NewGeneration$.append($S_i$)
\Else
\State $NewGeneration$.append($Population [i]$)
\EndIf
\EndFor
\State $Population  \gets NewGeneration$
\If{$\neg$ Improve($Population $)}
\State $life -=1$
\EndIf
\State $S_{best} \gets$ GetBestSolution($Population $)
 \EndWhile
\State \Return $S_{best}$
\EndFunction
\Function{Score}{$Candidate$}
   \State set tuned parameters according to $Candidate$
   \State $model \gets$TrainLearner()
   \State $result \gets$TestLearner($model$)   
   \State \Return$\mathit{Goal}(result)$  
\EndFunction
\Function{Extrapolate}{$old, pop, cr, f$}
  \State $a, b, c\gets threeOthers(pop,old)$  
  \State $newf \gets \emptyset$
  \For{$i=0 \to \mathit{np}-1$}
       \If{$cr < random()$}
         \State $newf$.append($old[i]$)
                \Else
                  \If{typeof($old[i]$) == bool}
                    \State $newf$.append(not $old[i]$)
         \Else
          \State $newf$.append(trim($i$,($a[i] + f * (b[i] - c[i]$)))) 
         \EndIf
       \EndIf
  \EndFor
 \State \Return $newf$
\EndFunction
        \end{algorithmic} 
\caption{Pseudocode for DE with Early Termination}
\label{alg:DE}
\end{algorithm}

@@

Our defect data comes from the PROMISE repository (http://openscience.us/repo/defect)
and pertains to 
open source Java systems defined in terms of \tab{ck}:  {\it ant}, {\it camel}, {\it ivy}, {\it jedit}, {\it log4j}, {\it lucene},
{\it poi}, {\it synapse}, {\it velocity} and {\it xerces}.

@@

\subsection{RQ2:  Does Tuning Change a Learner's Ranking ?}\label{sect:rank}
Researchers often use performance criteria to assert that one learner is better than 
another~\cite{lessmann2008benchmarking,me07b,hall11}. For example:
\be
\item
Lessmann et al.~\cite{lessmann2008benchmarking} conclude that
Random Forest is considered to be statistically 
better than CART. 
\item
Also, in Hall et al.'s   systematic literature review\cite{hall11}, it is argued
that defect predictors based on simple 
modeling techniques such as LR perform better than ``complicated'' techniques such as Random Forest.
To explain that comment, we note that
by three measures,
Random Forest
is more complicated than LR:
\be
\item
CART builds one model
while Random Forest builds many models. 
\item
LR is just
a model construction tool while Random Forest needs both
a tool to construct its forest {\em and} a second tool
to  infer some conclusion from all the members of that forest.
\item
the LR model can be printed in a few lines while the multiple
models learned by Random
Forest model would take up multiple pages of output.
\ee
\ee
Given tuning, how stable are these  conclusions?
Before answering this issue, we digress for two comments.

@@

@@

%%%%evaluations for pre and F %%%%%%
\begin{table}[!ht]
\renewcommand{\baselinestretch}{0.75}
\scriptsize
\centering
  \begin{tabular}{r|c c|c c|c c }
    Datasets & \multicolumn{2}{c|}{Tuned\_Where}  & \multicolumn{2}{c|}{Tuned\_CART} &  \multicolumn{2}{c}{Tuned\_RanFst}  \\
    \hline
    & precision  & F  & precision  & F & precision  & F  \\
    \hline
    antV0 & 50  &50 & 60&50  &60&  70 \\
    antV1 & 60  &60 &50& 50  &60&  60 \\
    antV2 & 70  &90& 50&60  &60& 120 \\
    camelV0 & 70  &50& 70 &80   &110&  70 \\
    camelV1 & 60  &60&60&110   &70&  70\\
    ivy & 60 &60&  60&60   &60&  60 \\
    jeditV0 & 80  &80&80& 60   &90&  60 \\
    jeditV1 & 60  &70&80&  70   &80& 70 \\
    jeditV2 & 90  &80&60&70  &110&  80 \\
    log4j & 50 &70& 50&50   &80&  50 \\
    lucene & 80 &60& 70&60  &60&  70 \\
    poiV0 & 60  &60& 70 &60  &130&  80 \\
    poiV1 & 50 &50& 70 &50   &50&  110 \\
    synapse & 70  &60&50&  60   &50&  90 \\
    velocity & 60  &60&50& 60   &100&  60 \\
    xercesV0 & 60  &80&80& 60   &70& 80 \\
    xercesV1 & 80  &80&60& 60   &50&  80 \\
  \end{tabular}
  \caption{Number of evaluations for tuned learners, optimizing for precision and F-Measure.}
\label{tab:eval}
\end{table}

%%%%runtime for prec and F %%%%%%
\begin{table*}[!ht]
\renewcommand{\baselinestretch}{0.75}
\scriptsize
\centering
  \begin{tabular}{r|c c|c c|c c|c c|c c|c c }
    Datasets & \multicolumn{2}{c|}{Tuned\_Where} & \multicolumn{2}{c|}{Naive\_Where} & \multicolumn{2}{c|}{Tuned\_CART} & \multicolumn{2}{c|}{Naive\_CART} & \multicolumn{2}{c|}{Tuned\_RanFst} & \multicolumn{2}{c}{Naive\_RanFst}\\
    \hline
     & precision  & F    &  precision  & F  & precision  & F & precision  & F & precision  & F & precision  & F\\
    \hline
    antV0 & 95.47  &93.58  & 1.65 &1.39&  5.08 &3.52& 0.08 &0.08& 9.78 &9.89& 0.20 &0.17\\
    antV1 & 224.67 &186.95 & 3.03 &3.18&  6.52 &6.18& 0.12 &0.12&  14.13 &13.39& 0.25 &0.25\\
    antV2 & 644.99 &654.34 & 8.24 &8.08&  9.00 &8.79& 0.24 &0.18&  16.75 &27.56& 0.44&0.36\\
    camelV0 & 690.62 & 543.28 & 7.93 &9.65&  12.68 &17.00& 0.24 &0.28&  28.49 &22.52& 0.34&0.41\\
    camelV1 & 1596.77 & 1808.03 & 23.56 & 26.98 &  17.13 &31.92& 0.27 &0.28&  33.96 &37.00& 0.77&0.85\\
    ivy & 66.69 & 74.50 & 0.97 &1.18 &  4.26 &4.72& 0.07 &0.08& 8.89 &10.39& 0.19&0.21\\
    jeditV0 & 459.30&518.47 & 5.33 & 6.11 &  8.69 &7.9& 0.11 &0.10& 18.40 &14.32& 0.32&0.37\\
    jeditV1 & 421.56&576.29 & 6.59 & 6.89 &  9.05 &8.13& 0.12 &0.10& 17.93 &17.42& 0.36&0.34\\
    jeditV2 & 595.56&657.59 & 6.88 & 7.93 &  7.90 &10.34& 0.14 &0.15& 27.34 &20.20& 0.38&0.40\\
    log4j & 76.09 &123.48 & 1.33 & 1.59 &  2.60 &2.92& 0.06 &0.08& 9.69 &7.67& 0.15&0.17\\
    lucene & 236.45& 219.02& 2.60 & 3.68 &  6.07 &6.89 & 0.10 &0.12& 9.77 &13.06& 0.25&0.35\\
    poiV0 & 263.12 &314.53 & 3.92 & 4.82 &  7.42 &7.80& 0.09 &0.10& 25.86 &19.29& 0.28&0.32\\
    poiV1 & 398.33 & 446.05 & 6.94 & 7.55 &  9.31 &7.62& 0.13 &0.14& 12.67 &27.23& 0.29&0.36\\
    synapse & 144.09 & 138.75 & 1.85 & 1.83 &  3.88 &4.87& 0.07 &0.08& 8.13 &13.29& 0.19&0.17\\
    velocity & 184.10 &211.88 & 2.68 & 3.13 &  4.27 &5.51& 0.07 &0.10& 15.18 &11.58& 0.21&0.27\\
    xercesV0 & 136.87 & 178.49 & 1.98 & 2.02 &  9.17 &7.47& 0.10 &0.11& 14.17 &17.31& 0.22&0.28\\
    xercesV1 & 1173.92 & 1370.89 & 12.78 &14.42 &  10.47 &11.07 & 0.16 &0.19& 18.27 &25.27& 0.40&0.46\\
  \end{tabular}
  \caption{Runtime for tuned and default learners(in sec), optimizing for precision and F-Measure.}
  \label{tab:runtime}
\end{table*}

The number of evaluations and runtime used by our optimizers are shown in \tab{eval}
and \tab{runtime}.
WHERE's runtime are slower than CART and Random Forest since WHERE has yet to benefit from decades
of implementation experience with these older algorithms. For example, SciKitLearn's  CART and Random Forest
 make extensive use of an underlying C library whereas WHERE is a purely interpreted Python.

Looking over \tab{eval}, the general pattern is that 50 to 80 evaluations suffice for finding the tuning
improvements reported in this paper. 
50 to 80 evaluations are  much fewer than our pre-experimental intuition.
Prior to this paper, the authors have conducted numerous explorations of evolutionary algorithms
for search-based SE applications~\cite{krall15,krall15:hm,fea02a,me07f,Green}. Based
on that work, our expectations were that non-parametric evolutionary optimization would
take thousands, if not millions, of evaluations of candidate tunings. This turned out not
to be that case. By comparing the runtime of tuned and default learners shown in \tab{runtime}, we notice
that the actual tuning time for most data sets is not extremely long.

@@

 In \fig{features}, we show how tuning selects the optimal values for tuned parameters. For space limitation, only four parameters from WHERE learner are selected as representatives and all the others can be found in our online support documents (https://goo.gl/aHQKtU).
 Note that
 the tunings learned were different in different data sets and for different goals.
Also, the tunings learned by DE
were often very different to the default (the default values for {\em threshold}, {\em infoPrune}, {\em min\_Size} and {\em wriggle} are $0.5$, $0.33$, $0.5$ and $0.2$, respectively). That is, to achieve the performance improvements seen in the paper,
the default tuning parameters required a wide range of adjustments.

@@

% \clearpage
% \pagenumbering{roman}
% \setcounter{page}{1} 
% \section*{Reply to Reviews}

% Thank your for your comments. The typos listed by reviewers
% have been fixed and the remainder of the paper has been given
% a careful proof read.

% The reviewers raised certain issues which we respond  to as follows.

% {\em I am concerned about the reproduction of work from elsewhere. Although the relevant paper and book are cited in section 2.3 ([27, 28]) the amount of material that is reproduced is surprisingly large. I am also unsure what "presenting some new results from Rahman et al. [28]" actually means. It would obviously be wrong to present other researchers' results as if newly discovered but I doubt that this is the intention. Indeed, on checking the other paper, there is no obvious overlap. This needs to be clarified.}
